---
title: "capstone1"
output: 
  md_document:
    variant: markdown_github

---

##Introduction
Children learn, adapt, and take on behaviors that are taught from the many influences that he/she may encounter. The number of influences that a person has is not limited to the family members in one’s house hold. A child’s many influence can include, but are not limited to, immediate family, friends, school, neighbors, celebrities, and the biggest of all the media. The media would be considered as all forms of communication outlets like the internet, social media, news media, print media, and in specific to this project video viewing and electronic games. The media can offer good and bad influences for all viewers watching it. It is a parent’s duty to protect the child from negative influences as much as possible to ensure healthy growth and implement good behavior.
	The media can introduce children to early exposure of media violence. Early exposure to media violence is a contributing factor to the development of aggression. The aggression then could lead to aggressive and violent behavior in children. Children are increasingly becoming heavy media consumers. Research indicates that much of the media directed at children contains violent content. Early exposure to media violence could place both male and female children at risk for the development of aggressive and violent behavior in adulthood. Children that watch shows that have violence that is realistic, consistently repeated and unpunished, are more likely to imitate what they see. Children with emotional, behavioral, learning or impulse control problems may be more easily influenced by video aggressiveness. The impact of TV violence may show immediately in the child's behavior or may surface years later. Due to the media presenting violence at a very high level, scientist have found ways parents can protect children from excessive media violence.   Each day there are multiple things and people that a child would observe where behaviors can be learned from. These multiple things would be considered as the factors of independent variables that affect behavior in children. In this project, the factors that are tested that affect behavior in children are electronics, siblings (older and younger), parental approach, and nutrition. These factors tend to be the major affects in children because these typically are things and people that the child would encounter every day. The examples set by adults, older siblings and children are the most powerful influences shaping a child's behavior and personality. Electronics a nutrition would be the next major influences on children behavior because electronic offer exposure to all types of human behavior from the negative to positive behavior. The elements or ingredients of a child's diet like thiamin treatment reverses aggressiveness in thiamin-deficient children, and sucrose and additives are suspected to induce hyperactivity.
  
  Of all types of behavior a child can display, the focus of the project would be on aggressive behavior. The level of aggressiveness would be determined according to the different circumstances of the household that the child lives in. it is know that all people's circumstances are different or have their own uniqueness. At a general level, realistic scenarios or cases that are close to what a child's family can fit in are projected according to the different levels of the project. Being that children would be the main focus, parents would be the audience of this project. 
The goal of the project would be to predict the level of aggression with these given levels according to each factor or independent variable. The variables are lised below:


Dependent Variable: 
Aggression - aggression level seen in behavior

Independent Variables:
Videos - the amount of time spent viewing videos
Electronic Games - the amount of time spent playing electronic games
Siblings - amount of aggression seen in older siblings
Nutrition - detremined healthy or unhealthy diet
Parental Approach - how the parents implement parenting skills effectively


 
The necessary libraries needed for the project has been installed and are loaded here.

```{r, warning=FALSE, message=FALSE}
#library(plyr)
#library(dplyr)
#library(tidyr)
#library(readxl)
library(tidyverse)
#library(ggplot2)
#library(caret)
#library(RCurl) #To read the csv file from GitHub directly
library(rmarkdown)
```

All of the packages are loaded.  Therefore, we load the data into R and call the data frame dta.


```{r}
dta <- read.csv("CAclean.csv")
```


An extra variable was included in the data frame after the data was read.  Thus, the variable was removed.
```{r}
dta<-select(dta, -c(X))
```

##Data Wrangligling

The data set was provided in a nearly cleaned manner.  However, multiple data wrangling techniques were performed on the data frame to transom it into a document that can be analyzed.  These techniques include removing unnecessary variables, changing the names of some columns, adding variables, and including functions for calculation. 

Changing Columns Names:  There are six variables included in the data frame, and all variables underwent a name change.  When the data was imported into R Studio, some variables names were too long, and others did not accurately describe the particular data.  Therefore, the names of the variables were changed to shorter, concise names that accurately describe the data in the column. AggUk was changed to Aggression. VideoUK was changed to Videos. ElectUK was changed to Electronig games. SibUK was changed to Siblings. NutUK was changed to Nutrition. ParUK was changed to Parental Approach. The names of the columns had to become broader to fit the data of both the United States and the United Kingdom. 

Adding Variables and Data:Two variables were necessary for building the model.  A factor column and a numerical column that UK or US for each child. The numeric values will allow the United States and the United Kingdom to be compared for the charts and graphs. The United States was given the value of zero, while the United Kingdom was given the value of one.   




##Exploratory Data Analysis
Now that the data have been ingested it is time to begin the exploration.  The first thing to do is to have a quick look at our data.  For that, the `str()` function and the more human-readable (and intuitive), `glimpse()` function. is used.

```{r}
glimpse(dta)
```


As expected, there are 667 rows with 8 variables inside of the data set.  Additionally, the type of variables is apparent from this view.  For example, UK.US` is a factor variable.   A factor, instead of a number seems more logical in this case.  Necessary changes will be made to the variable if the exploratory analysis dictates.  For now, we proceed with the intended analysis.  Now let's take a look of the structure of the data set to see what additional information is available.  
```{r}
str(dta)
```

Evaluating the output for each code, it is easy to determine that the codes provide almost identical information.  


```{r}
library(tidyverse)
dta %>% group_by(Aggression) %>% count() %>% arrange(desc(n))
```


The aggression level with the largest amount seen in children was 0.040.  This number shows that there is some level of aggression in the child.


```{r}
agg_count <- dta %>% group_by(Aggression) %>% count() %>% filter(n > 5)
agg_count
```


 A comparison among aggression in the two countries will provide pertinent information. The results will be displayed as a bar graph. 

```{r}
ggplot(dta %>% filter(Aggression %in% unique(agg_count$Aggression)),
       aes(x = UK_US, fill = Aggression)) + geom_bar(position = 'dodge')
```